Enhancing the prediction of sales for newly published books by means of computational intelligence methods

### 2016-05-21 paper.tex
Working on the explanation of the trees
This is one of the main results of our paper and it was dispatched with a couple of lookie here, this graphic and some paragraphs. What we have to do in this paper is to actually look at the factors that make a book sell, and that is contained within those trees so I have added a good amount of explanation to them.

Refs #39 #25

As it can be seen in this table, \texttt{total\_sales} is the output
variable, so it is the objective of the prediction methods. {\tt
  total_sales} are a variable percentage of the {\tt print_run}, and
thus there is obviously some correlation between them, which might
prevent us from using it to predict sales; it is quite obvious that
just increasing the print run, by itself, will not boost
sales. However, several of the other variables include implicitly a
certain size of the print run; for instance, you have to actually
print books to distribute them as novelty and at least one, for every
point of sale, so including the print run makes sense to get more
accurate estimates of the total sales that you are going to obtain out
of the total print run and check its influence on the overall sales
volume. Once this is included in the model and also as part of the
tree nodes, changing the print run might give the publisher an idea of
how many books are going to be actually sold out of that print run,
and taking that into account decide on the actual number of books that
are going to be printed. Just to be clear, printing book does not
actually sell them, but it is obviously necessary to print books
before you sell them, and once you have to do it, the publisher might
make decisions on printing a few more and check, with this model, what
would be the influence in sales if, for instance, you decide to send
more books to every point of sale or distribute more in novelty. We
will come back to this choice later on when examining the resulting
models.

@@

In this section we show the output of one of the models built during the regression, namely we present two of the decision trees generated by M5P method applied to the data of the publisher that sells the most. These aim to demonstrate the value of the proposal to generate a useful decision-aid tool for an expert (publisher) to determine what sales can be expected for the different values of the input variables (including the print-run).

First, Figure \ref{fig:decision_tree_best_seller} depicts the decision tree for the publishing company UDL929. The leaves correspond to linear regression models that can be computed (easily) to get the final predicted value for book sales. 
The second decision tree (Figure \ref{fig:decision_tree_best_seller_fs}) corresponds to the same publishing company, so the same data were used, but the RAE feature selection method was applied. Thus, the main difference in the simplification of the regression models in the second tree and the remove of \texttt{print\_run} as part of the decision process after the feature selection.

Examining both trees, it can be noticed that \texttt{tot\_points\_sale} variable appears twice in the trees, being also the root. This indicates that the total number of points of sale has a big influence on the decision models and also a major role in forecasting the total sales. This is an interesting clue for the experts, who could check the expected impact of different number of selling points
According to the trees, \texttt{distrib\_novelty} has also a big repercussion, appearing at the second level of the top of the trees.

An important fact that these tree models reveal is that marketing variables, such as the total number of points of sale and the number of units distributed as novelty have higher influence on the prediction than the selling price of the book, which could be, initially, considered as one of the most important parameters to determine sales. This interpretation leads us to suggest that companies could probably improve the accuracy of their sales forecasting models by incorporating more marketing related variables in their models.

\begin{figure}[!ht] 
\begin{center}
  \epsfig{file=decision_tree_no_fs.eps,width=13cm}
\caption{Decision Tree built by the M5P method for the best seller publishing company with all the pre-sales features.}
\label{fig:decision_tree_best_seller}
\end{center}
\end{figure}

\begin{figure}[!ht] 
\begin{center}
  \epsfig{file=decision_tree_fs.eps,width=13cm}
\caption{Decision Tree built by the M5P method for the best seller publishing company with the features selected by the ReliefAttributeEval method.}
\label{fig:decision_tree_best_seller_fs}
\end{center}
\end{figure}

>>>>

As it can be seen in this table, \texttt{total\_sales} is the output
variable, so it is the objective of the prediction methods. {\tt
  total\_sales} are a variable percentage of the {\tt print\_run}, and
thus there is obviously some correlation between them, which might
prevent us from using it to predict sales; it is quite obvious that
just increasing the print run, by itself, will not boost
sales. However, several of the other variables include implicitly a
certain size of the print run; for instance, you have to actually
print books to distribute them as novelty and at least one, for every
point of sale, so including the print run makes sense to get more
accurate estimates of the total sales that you are going to obtain out
of the total print run and check its influence on the overall sales
volume. Once this is included in the model and also as part of the
tree nodes, changing the print run might give the publisher an idea of
how many books are going to be actually sold out of that print run,
and taking that into account decide on the actual number of books that
are going to be printed. Just to be clear, printing book does not
actually sell them, but it is obviously necessary to print books
before you sell them, and once you have to do it, the publisher might
make decisions on printing a few more and check, with this model, what
would be the influence in sales if, for instance, you decide to send
more books to every point of sale or distribute more in novelty. We
will come back to this choice later on when examining the resulting
models.

@@

What is, then, the model obtained with our method? What factors have
the biggest influence in sales and how can publisher use them to gauge
sales of a new book? In order to do that, we have examined the models
built during the regression, namely we present two of the decision
trees generated by M5P method applied to the data of the publisher
that sells the most. By looking at these trees, the publisher should
be able to decide how many actual books should be printed to put them
for sale, which was the initial objective of our research.

The model obtained for the publishing company UDL929 is shown in
Figure \ref{fig:decision_tree_best_seller}, which was the company that
sold the most books in our data set. It makes sense to use data for a
single company, since book sales for all companies is only available
if you buy it; a single company, however, will have complete data on
its own sales. This company is also representative because, being the
one that sells the most, will have the biggest range of sales for
individual books. In this figure, tree {\em leaves} correspond to
linear regression models that can be easily computed to obtain the
final predicted value for book sales.
The second decision tree (Figure
\ref{fig:decision_tree_best_seller_fs}) corresponds to the same
publishing company, so the same data were used, but the RAE feature
selection method was applied. This simplification, besides making the
tree easier to understand, reduces training time to a fraction of what
was needed before it.

Examining both trees, it can be noticed that
\texttt{tot\_points\_sale} variable appears twice in the trees, being
also the root. This indicates that the total number of points of sale
has the biggest influence on the decision models and also a major role in
forecasting the total sales, implying that buyers have to actually
{\em see}, maybe browse the book, before buying it. Books have to
physically {\em be} in the point of sale, so this number gives a lower
bound to the number of books that have to be printed; a stack of book
will have to be present in each point of sale to attain that effect. 
 This is an interesting clue for the
experts, who could check the expected impact of different number of
selling points. This feature also shows that there are different
regimes depending on the number of points of sale, with best sellers
being those that are distributed to more than roughly 500 points.

The second {\em splitting} feature, number of books distributed as a
novelty, gives the publisher a hint on how many books have to be sent
to bookstores to stack on the {\em novelty} table. You need to sell
roughly 5, the result of dividing 2807.5 y 550.5, to have a lower
bound in sales at 1006, which is negatively impacted by the retail
price and the number of points of sale in the first year of
distribution. In all cases, the last number in the formula featured in
the box gives you a baseline for sales, which the rest of the
variables impacting positively or negatively. A whole treatise could
be written on this decision tree but, in this case, let us look a bit
more closely at this box and try and interpret it. For instance, price
in this case will impact negatively on sales. For a typical retail
price of a trade paperback of 12€, we would be looking at negative
sales (if we disregard the rest of the variables) , which in this case
will indicate that these are the kind of books, maybe paperbacks, that
sell for a low price of around 5€. You also have to look at the
relation between the factor for total points of sale and the total
point of sales the first year. The fact that this last one is bigger
indicates that the publisher will have to {\em increase} the number of
points of sale after the first year, distributing it more shops at a
more capilar level; the number of points of sale will have to be
increased by 10\% to achieve a bit more than the baseline. All in all,
the conclusion with this part of the tree is that you will have to
print around 3000 books to sell a quantity that will start at 1000,
one third of the print run, and changes depending on a number of
factors, including price.

An important fact that these tree models reveal is that marketing
variables, such as the total number of points of sale and the number
of units distributed as novelty have higher influence on the
prediction than the selling price of the book, which could be,
initially, considered as one of the most important parameters to
determine sales. This interpretation leads us to suggest that
companies could probably improve the accuracy of their sales
forecasting models by incorporating more marketing related variables
in their models. This does not mean that the retail price is not a
factor; it appears in all models, but it will be a secondary factor
that will have to be decided once the main ones are settled.

\begin{figure}[!ht] 
\begin{center}
  \epsfig{file=decision_tree_no_fs.eps,width=13cm}
\caption{Decision tree built by the M5P method for the best seller publishing company including all pre-sales features.}
\label{fig:decision_tree_best_seller}
\end{center}
\end{figure}

\begin{figure}[!ht] 
\begin{center}
  \epsfig{file=decision_tree_fs.eps,width=13cm}
\caption{Decision tree built by the M5P method for the best seller
  publishing company using  the subset of features selected by the
  {\sf ReliefAttributeEval} method.} 
\label{fig:decision_tree_best_seller_fs}
\end{center}
\end{figure}
